The Decreasing Orbital Period of WASP-4\lowercase{b}

### 2018-12-06 paper/ms.tex
improvements to the text

\begin{abstract}
  %TODO: final numbers
  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
  with previous studies, we show that the transit times of WASP-4b are
  incompatible with a constant orbital period. 
  In particular, the transits seem to have arrived $77.8 \pm 10.7$
  seconds early, and the period appears to be shrinking by
  $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
  From TESS observations of WASP-6b, WASP-18b, and WASP-46b, we show
  that a systematic offset between the TESS time stamps and the barycentric
  reference sufficient to explain WASP-4b is ruled out at 6.3$\sigma$.
  If the timing variations are astrophysical,
  major contributors to the period change could include both
  apsidal precession and tidal decay.
  The Doppler shift from WASP-4's acceleration towards us could
  account for at most one quarter of the observed period change
  (at $2\sigma$).
  Further transit and occultation studies will help confirm the
  reality of the timing variation, and eventually determine its cause.
\end{abstract}

@@

Long-term monitoring of hot Jupiter transit and occultation times should
eventually reveal two distinct processes.  First, whenever the summed
spin and orbital angular momenta in a hot Jupiter system are below a
particular critical value, the system will be unstable to tidal decay
\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
satisfy this condition; their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay is
uncertain, and depends on how friction dissipates energy carried by
gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
we could directly measure orbital decay rates, it would inform our
understanding of tidal dissipation.  This is important because the
dissipation rates determine the fate of the system.  During its
inspiral, if the planet fills its Roche lobe below the stellar surface,
the drag force from the stellar envelope accelerates the inspiral,
leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above above the stellar surface, then it
could either be dynamically disrupted into an accretion disk or it could
slowly transfer mass to the star
\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
Further afield, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

The second process that long-term timing studies should reveal is
rotation of the orbital ellipse within the orbital plane (apsidal
precession).  This effect has been studied extensively in eclipsing
binaries \citep[{\it e.g.},][]{
schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}, and has been
explored theoretically for transiting planets
\citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
However, apsidal precession has yet to be directly observed in hot
Jupiters, and the orbits would need non-zero eccentricity for it to be
detectable.  Though the eccentricities of hot Jupiters are expected to
be small, there are many mechanisms that could pump them away from
circularity (see \S~\ref{sec:apsidal_precession}, and
\citealt{bailey_understanding_2019}).  For hot Jupiters, the apsidal
precession is usually dominated by the quadrupolar distortion of the
planet by the star's tidal force \citep{ragozzine_probing_2009}.  If we
measured apsidal precession in a hot Jupiter system, we could thus use
the observed precession rate to compute the planet's Love number, giving
a third parameter along with the mass and radius with which to describe
the planet's interior structure \citep[{\it e.g.},][who performed a
similar procedure for HAT-P-13b]{batygin_determination_2009}.  Measuring
small eccentricities and understanding their origin would also help us
understand the dynamical histories and interior structures of hot
Jupiters.  \citep[{\it
e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}

While numerous indirect studies have called attention to
population-level effects of tidal decay
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
the most convincing direct evidence to date for either orbital decay or
apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence, and
measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
\citet{patra_2017} presented new transit and occultation times for the
system, and found $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.
Assuming that the period change was caused entirely by tidal decay,
\citet{patra_2017} found values for the stellar quality factor of
$Q_\star' \approx 2\times10^5$.  This level of dissipation cannot be
explained by standard equilibrium tide
models~\citep{penev_tidal_2011,ogilvie_tidal_2014}.
\citet{weinberg_tidal_2017} found though that if WASP-12 were a
subgiant, tidal dissipation due to a dynamical tide might be sufficient
to produce the required level of friction.
\cite{bailey_understanding_2019} concurred, but noted that the
observational constraints on WASP-12 itself favor a main-sequence star,
not a subgiant.  More data are needed to determine the cause of the
timing variations in WASP-12b.

This work focuses on WASP-4b, which is also a plausible candidate for
orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$, and
mass $1.29 \pm 0.10\ M_{\rm Jup}$ \citep{southworth_homogeneous_2011}.
It has an orbital period of 1.34 days, orbits its host star at a
distance of $a/R_\star = 5.41$, and has been frequently observed over
the past decade \citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The
orbit is consistent with circular \citep[][but see
\S~\ref{sec:apsidal_precession}]{beerer_secondary_2011,husnoo_observational_2012,bonomo_gaps_2017}.
The host star is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
$M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
\citep{southworth_homogeneous_2011,petrucci_no_2013,doyle_accurate_2013}.

@@

\subsection{Observations and cleaning}
WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input Catalog as TIC 402026209
\citep{stassun_TIC_2018}.

The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
array every 2 minutes.  It was included on the ``short-cadence'' target
list thanks to the Guest Investigator programs of J.\ Southworth and S.\
Kane (G011112 and G011183 respectively).  After being downlinked through
the Deep Space Network, the data were processed through the Science
Processing Operations Center (SPOC) pipeline at NASA
Ames~\citep{jenkins_tess_2016}.  One important step in this processing
was converting from spacecraft time to barycentric time.  SPOC then
reports timestamps in ``BTJD'', where ${\rm BTJD} = {\rm BJD} -
2457000$, BJD is the barycentric julian date, and all times are in the
TDB reference (see \citealt{urban_explanatory_2012} for explanations of
time systems).  The lightcurves were then vetted and released by the MIT
TESS Science Office to the Mikulski Archive for Space
Telescopes~\citep{ricker_tess_alerts_2018}.

@@

Using the times, flux measurements, and errors determined above, we ran
Box Least Squares (BLS) to estimate the transit epoch, period, and
duration \citep{kovacs_box-fitting_2002}.  From the BLS parameters, we
isolated each transit to within $\pm$10 transit durations of its
midtime.  In each transit window, we then simultaneously fit for a model
transit, plus a local linear trend.  Our transit model uses the analytic
formulae calculated by \citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.

@@

% TODO: final numbers
To check that the measured uncertainties are plausible, we computed the
reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
Examining the residuals by eye showed that the error variance was being
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.73$, forcing a reduced $\chi^2$ of unity.  This reduced the mean
transit time uncertainty from $\sim$30 seconds per transit to $\sim$22
seconds per transit.

Figure~\ref{fig:lightcurves} shows the light curves, best-fit transit
models, and residuals.  Table~1 reports the mid-transit times and
their uncertainties, and also includes the previously reported times
analyzed in \S~\ref{sec:timing}.

@@

Binning the residuals to 1-hour windows and taking the median absolute
deviation, we measure a MAD of the residual lightcurves of $594\,{\rm
ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
\citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
noise\footnote{\url{github.com/lgbouma/tnm}} of $410\,{\rm ppm}\,{\rm
hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which was used in the
\citet{Sullivan_2015} planet detection simulations.  The predicted level
of zodiacal noise seems to have been an overestimate, given the observed
RMS of WASP-4.  A more detailed assessment of TESS's photometric
performance is beyond the scope of this work.

@@

\subsection{Times}
\label{subsec:times}

@@

Table~1 shows all the transit times used in our analysis.  We included
data from peer-reviewed literature for which the analysis was based on
observations of a single, complete transit, and for which the midpoint
was fit as a free parameter. We also required that the time system be
clearly documented.

@@

We use the homogeneous \citet{hoyer_tramos_2013} timing study as our
starting point.  We took their BJD$_{\rm TT}$ times to be equivalent to
times in BJD$_{\rm TDB}$, since the two time systems differ periodically
by only milliseconds \citep{urban_explanatory_2012}.  We verified that
their times, as well as all others, were on the BJD$_{\rm TDB}$ time
standard using the calculator provided
by~\citet{eastman_achieving_2010}.  We set the zero-point epoch to be as
close as possible to the average of the transit times, weighted as the
inverse midtime variance. This minimizes the covariance between the
transit epoch and the period \citep{gibson_gemini_2013}.

Important contributions to the timing dataset are as follows.  The
earliest epoch in our fit is from the EulerCam lightcurve observed by
\citet{wilson_wasp-4b_2008}.  The second, more precise, epoch comes from
$z$-band photometry acquired by \citet{gillon_improved_2009} at the VLT
with FORS2.  Subsequent epochs came from
\citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
the GPS clock time in UT reference. They then converted this to BJD in
the TDB reference for their reported midtimes.},
\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
\citet{nikolov_wasp-4b_2012}.
%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS
%clock time in UT reference. They then converted this to BJD in the
%TDB reference for their reported midtimes.}, and reported transit
%midtimes precise to 6~seconds.  \citet{dragomir_terms_2011} updated
%the ephemeris using two transits measured with the CTIO 0.9-m and
%1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
%transits from the Baade telescope at Las Campanas\footnote{
%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
%to show that the star's rotation axis is nearly aligned with the
%planet's orbital axis, in agreement with the Rossiter-McLaughlin
%measurement by \citet{triaud_spin-orbit_2010}.}.
%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
%the MPG/EOS-2.2~m at La Silla.
\citet{hoyer_tramos_2013} reported nine transit observations taken with
the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
SOI on the SOAR 4.2-m telescope.  \citet{hoyer_tramos_2013} also
homogeneously redetermined the transit times from all previous studies.
Though their latest two epochs fell slightly below the prediction of a
linear ephemeris (Panel C of their Figure~6), they did not find a need
to fit a quadratic function to the O-C values.

After the study by \citet{hoyer_tramos_2013},
\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
using HST's WFC3 in both transit and secondary eclipse.  We adopt only
their transit time, since the epoch of occultation was not fit as a free
parameter.  \citet{huitson_gemini_2017} acquired optical transmission
spectra with the 8.1-m Gemini South telescope over 2011 to 2014, one
transit per season.  The per-point standard deviation of their
lightcurves is a few hundred parts per million, less than a factor of 3
away from the photon noise limit.  The uncertainties on their reported
transit times~--~on average 5.6~seconds~--~ are small, but possible
given the quality of their data~\citep[cf.][]{carter_analytic_2008}.

There have been other studies of WASP-4b for which the transit timing
data are either hetereogenous, not available, or not useable.  We
omitted the timing measurements from
\citet{southworth_high-precision_2009}, since there were technical
problems with the computer clock at the time of
observation~\citep{nikolov_wasp-4b_2012}.  The timing study
by~\citet{petrucci_no_2013} ruled out transit timing variations larger
than 54~seconds over the previous 5 years, but did not report transit
times.  \citet{baluev_benchmarking_2015} collected transit times for
many hot Jupiters, including WASP-4b, and included times from the
Exoplanet Transit Database
(ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
\citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
sources, their timestamps are less clearly documented, and the times
are thus more prone to systematic errors, we omit them from
consideration.
%We consider the effects of including ETD data in
%Appendix~\ref{sec:verify_archival_times}.
\citet{may_mopss_2018} obtained optical spectra using IMCAS on
Magellan, and found a flat transmission spectrum, in agreement with
the results reported by \citet{huitson_gemini_2017}.  They did not
report transit times.

There are fewer available occultation times.
\citet{beerer_secondary_2011} observed two occultations of WASP-4b using
warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
\citet{caceres_ground-based_2011} detected an occultation from the
ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
specifying the time standard.  We correspondingly added $69.184$ seconds
of uncertainty to their reported errors, in quadrature.  Finally, as
previously mentioned, \citet{ranjan_atmospheric_2014} acquired
occultation data with HST but kept the epoch as a fixed parameter.
\citet{zhou_secondary_2015} performed a similar analysis with
occultation data from the Anglo-Australian Telescope, focusing on the
eclipse depth, but also fitting for $e\cos\omega$.  We converted their
$e\cos\omega$ results into a midtime, using their ephemeris and the
standard formula \citep[{\it
e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
\begin{equation}
  t_{\rm occ}(E) =
  t_0 +  P E  +
  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
  \label{eq:occultation_time}
\end{equation}
This gives us a total of 4 occultations.  Although their timing error
bars are large compared to the transits, we include them for
completeness, and because they can help in our modeling of the timing
variations.

@@

After collecting the literature times and measuring the TESS times, we
compared the observed time to the prediction based solely on the
literature times. The result is shown in
Figure~\ref{fig:arrived_early}. The TESS observations appear to have
arrived early.  One immediate concern was systematic errors in either
the archival or the TESS timestamps. The tests we used to mitigate the
latter possibility are described in Appendix~\ref{sec:verify_tess}.

@@

Assuming the error bars on the transit midtimes are gaussian and
independent, we then proceed by fitting each model to the timing data.
For the linear model, we took uniform priors over small windows around
the expected parameters.  For the quadratic model we did the same, and
took the quadratic term to have a uniform prior corresponding to
$
  Q_\star' \sim \mathcal{U}[10^3, 10^9],
$
for $Q_\star'$ the modified quality factor defined in
\S~\ref{sec:implications}.  For the precession model, we took the same
priors in epoch and sidereal period, and drew the remaining parameters
from wide uniform distributions.

%TODO: final numbers
Figure~\ref{fig:times} shows the residuals with respect to the linear
model.  The median MCMC linear fit has $\chi^2 = 149$ and 58 degrees of
freedom.  The median MCMC quadratic fit has $\chi^2 = 48.0$ and 57
degrees of freedom.  The median MCMC precession fit has $\chi^2 = 51.0$
and 55 degrees of freedom.

%TODO: final numbers
The difference in $\chi^2$ between the linear and quadratic fit
corresponds to $p \approx 10^{-23}$. In other words, assuming the linear
model is true, there is a $10^{-23}$ probability that simply by chance
we would see data that favors the quadratic model more strongly than the
observations.  A more likely possibility is some systematic error being
present in one of the most important sets of times: either the earliest
times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591, and
851), or the TESS times.  We have taken every precaution against
systematic offsets in the TESS dataset (Appendix~\ref{sec:verify_tess}),
and Huitson et al.\ stand by their reported times (Huitson, priv.
comm.).  Since the linear model provides a poor fit to the data, we
discard it from further consideration.

>>>>

\begin{abstract}
  %TODO: final numbers
  Combining data from the Transiting Exoplanet Survey Satellite (TESS)
  with previous studies, we show that the transit times of WASP-4b are
  incompatible with a constant orbital period. 
  In particular, the transits seem to have arrived $77.8 \pm 10.7$
  seconds early, and the period appears to be shrinking by
  $\dot{P}=-12.1 \pm 1.2$ milliseconds per year.   
  From TESS observations of WASPs-6b, -18b, and -46b, we show
  that a systematic offset between the TESS time stamps and the barycentric
  reference sufficient to explain WASP-4b is ruled out at 6.3$\sigma$.
  If the timing variations are astrophysical,
  major contributors to the period change could include both
  apsidal precession and tidal decay.
  The latter seems implausible on theoretical grounds, as it would
  require a stellar quality factor $Q_\star' \approx 3\times10^4$.
  The Doppler shift from WASP-4's acceleration towards us could
  account for at most one quarter of the observed period change
  (at $2\sigma$).
  Further transit and occultation studies will help confirm the
  reality of the timing variation, and eventually determine its cause.
\end{abstract}

@@

Long-term monitoring of hot Jupiter transit and occultation times should
eventually reveal two distinct processes.  First, whenever the summed
spin and orbital angular momenta in a hot Jupiter system are below a
critical value, the system will be unstable to tidal decay
\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot Jupiters
satisfy this condition; their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay is
uncertain, and depends on how friction dissipates energy carried by
gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
we could directly measure orbital decay rates, it would inform our
understanding of tidal dissipation.  This is important because the
dissipation rates determine the fate of the system.  During its
inspiral, if the planet fills its Roche lobe below the stellar surface,
the drag force from the stellar envelope accelerates the inspiral,
leading to a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above above the stellar surface, then it
could either be disrupted into an accretion disk or it could
slowly transfer mass to the star
\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
Further afield, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

The second process that long-term timing studies should reveal is
rotation of the orbital ellipse within the orbital plane (apsidal
precession).  This effect has been studied extensively in eclipsing
binaries \citep[{\it e.g.},][]{
schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}, and has been
explored theoretically for transiting planets
\citep{heyl_using_2007,pal_periastron_2008,jordan_observability_2008,ragozzine_probing_2009}.
However, apsidal precession has yet to be directly observed in hot
Jupiters, and the orbits would need non-zero eccentricity for it to be
detectable.  Though the eccentricities of hot Jupiters are expected to
be small, there are many mechanisms that could pump them away from
circularity (see \S~\ref{sec:apsidal_precession}, and
\citealt{bailey_understanding_2019}).  For hot Jupiters, the apsidal
precession is usually dominated by the quadrupolar distortion of the
planet by the star's tidal force \citep{ragozzine_probing_2009}.  The
observed precession rate in a hot Jupiter system can thus be used to
compute the planet's Love number, giving a third parameter along with
the mass and radius with which to describe the planet's interior
structure \citep[{\it e.g.},][who performed a similar procedure for
HAT-P-13b]{batygin_determination_2009}.  Measuring small
eccentricities and understanding their origin might also help us
understand the dynamical histories and inflated radii of hot Jupiters.
\citep[{\it
e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}

While numerous indirect studies have called attention to
population-level effects of tidal decay
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
the most convincing direct evidence to date for either orbital decay
or apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.
Using new timing data, \citet{patra_2017} found
$\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If
the period change were caused entirely by tidal decay, the implied
modified stellar quality factor would be $Q_\star' \approx 2\times10^5$.
Standard equilibrium tides do not predict this
level of dissipation \citep{penev_tidal_2011,ogilvie_tidal_2014}.  The
needed level of friction might come from dynamical tides, or
obliquity tides
\citep{weinberg_tidal_2017,millholland_obliquity_2018}.  The former is
more likely if WASP-12 is a subgiant, but the observational
constraints favor a main-sequence model
\citep{bailey_understanding_2019}.  More data are needed to understand
the WASP-12b timing variations.

This work focuses on WASP-4b, which is also a plausible candidate for
orbital decay.  The planet has radius $1.37 \pm 0.04\ R_{\rm Jup}$,
and mass $1.29 \pm 0.10\ M_{\rm Jup}$
\citep{southworth_homogeneous_2011}.  It has an orbital period of 1.34
days, orbits its host star at a distance of $a/R_\star = 5.41$, and
has been frequently observed over the past decade
\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  The orbit is
consistent with circular (but see \S~\ref{sec:apsidal_precession}),
and has a sky-projected obliquity consistent with zero
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.
The host is a G7V dwarf, with $R_\star = 0.92 \pm 0.13\ R_\odot$,
$M_\star = 0.92 \pm 0.07\ M_\odot$, $T_{\rm eff} = 5400 \pm 90\ {\rm
K}$, and an isochronal age of roughly $7\ {\rm Gyr}$
\citep{southworth_homogeneous_2011,petrucci_no_2013,doyle_accurate_2013}.

@@

\subsection{Observations}
WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input Catalog as TIC 402026209
\citep{stassun_TIC_2018}.

The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
array every 2 minutes.   After being downlinked through
the Deep Space Network, the data were processed through the Science
Processing Operations Center (SPOC) pipeline at NASA
Ames~\citep{jenkins_tess_2016}.  One important step in this processing
was converting from spacecraft time to barycentric time.  SPOC then
reports timestamps in ``BTJD'', where ${\rm BTJD} = {\rm BJD} -
2457000$, BJD is the barycentric julian date, and all times are in the
TDB reference (see \citealt{urban_explanatory_2012} for explanations of
time systems).  The lightcurves were then vetted and released by the MIT
TESS Science Office to the Mikulski Archive for Space
Telescopes~\citep{ricker_tess_alerts_2018}.

@@

Using the times, relative fluxes, and errors determined above, we ran
Box Least Squares (BLS) to estimate the transit epoch, period, and
duration \citep{kovacs_box-fitting_2002}.  Using these parameters, we
isolated each transit to within $\pm$10 transit durations of its
midtime.  In each transit window, we then simultaneously fit for a model
transit, plus a local linear trend.  Our transit model uses the analytic
formulae calculated by \citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.

@@

% TODO: final numbers
To verify that the measured uncertainties are plausible, we computed the
reduced $\chi^2$ for a linear ephemeris fit to the measured TESS transit
times.  We found that $\chi^2 = 8.4$, with 16 degrees of freedom.
Visually inspecting residuals showed that the error variance was being
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
midtime uncertainty for each transit from 30 to 22
seconds.

Figure~\ref{fig:lightcurves} shows the light curves, best-fit transit
models, and residuals.  Table~1 reports the mid-transit times and
their uncertainties.

@@

Binning the residuals to 1-hour windows and taking the median absolute
deviation, the residual lightcurves have a MAD of $594\,{\rm
ppm}\,{\rm hr}^{1/2}$.  WASP-4 has a TESS-band magnitude of $T=11.78$
\citep{stassun_TIC_2018}, corresponding to a predicted photon-counting
noise\footnote{\url{github.com/lgbouma/tnm}} of $410\,{\rm ppm}\,{\rm
hr}^{1/2}$.  At $T=11.78$, with a 9-pixel aperture read noise is
expected to contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015}
used in planet detection simulations.  The predicted level
of zodiacal noise seems to have been an overestimate, given the observed
RMS of WASP-4.  A more detailed assessment of TESS's photometric
performance is beyond the scope of this work.

@@

\subsection{Previous timing measurements}
\label{subsec:times}

@@

Table~1 shows the transit times used in our analysis.  We included
data from peer-reviewed literature for which the analysis was based on
observations of a single, complete transit, and for which the midpoint
was fit as a free parameter. We also required that the time system be
clearly documented.

@@

We used the homogeneous \citet{hoyer_tramos_2013} timing study as our
starting point.  
Though their latest two epochs fell slightly below the prediction of a
linear ephemeris (Panel C of their Figure~6), they did not find a need
to fit a quadratic function to the O-C values.
We verified that the \citet{hoyer_tramos_2013} times, as well as all
others, were on the BJD$_{\rm TDB}$ time
standard using the \citet{eastman_achieving_2010} calculator.
We set the zero-point epoch to be as
close as possible to the average of the transit times, weighted as the
inverse midtime variance. This minimizes the covariance between the
transit epoch and the period \citep{gibson_gemini_2013}.

Important contributions to the timing dataset are as follows.  The
earliest epoch in our fit is from EulerCam on the 1.2-m Euler telescope
\citep{wilson_wasp-4b_2008}.  The second epoch comes from
$z$-band photometry acquired by \citet{gillon_improved_2009} at the VLT
8.2-m with FORS2.  Subsequent transits came from
\citet{winn_transit_2009}\footnote{\citet{winn_transit_2009} recorded
the GPS clock time in UT reference. They then converted this to BJD in
the TDB reference for their reported midtimes.},
\citet{dragomir_terms_2011}, \citet{sanchis-ojeda_starspots_2011}, and
\citet{nikolov_wasp-4b_2012}, \citet{hoyer_tramos_2013}, and
\citet{ranjan_atmospheric_2014}.
%\citet{winn_transit_2009} measured two transits with the Baade 6.5-m
%at Las Campanas\footnote{\citet{winn_transit_2009} recorded the GPS
%clock time in UT reference. They then converted this to BJD in the
%TDB reference for their reported midtimes.}, and reported transit
%midtimes precise to 6~seconds.  \citet{dragomir_terms_2011} updated
%the ephemeris using two transits measured with the CTIO 0.9-m and
%1.0-m.  \citet{sanchis-ojeda_starspots_2011} reported four more
%transits from the Baade telescope at Las Campanas\footnote{
%\citet{sanchis-ojeda_starspots_2011} also used starspot occultations
%to show that the star's rotation axis is nearly aligned with the
%planet's orbital axis, in agreement with the Rossiter-McLaughlin
%measurement by \citet{triaud_spin-orbit_2010}.}.
%\citet{nikolov_wasp-4b_2012} observed 3 transits, in 4 independent
%photometry bands (Sloan {\it g'}, {\it r'}, {\it i'}, {\it z'}) with
%the MPG/EOS-2.2~m at La Silla.
%\citet{hoyer_tramos_2013} reported nine transit observations taken with
%the the Y4KCAM on the SMARTS 1-m telescope, and another three with the
%SOI on the SOAR 4.2-m telescope.
%\citet{hoyer_tramos_2013} also
%homogeneously redetermined the transit times from all previous studies.
%After the study by \citet{hoyer_tramos_2013},
%\citet{ranjan_atmospheric_2014} acquired near-IR spectra of WASP-4b
%using HST's WFC3 in both transit and secondary eclipse.  We adopt only
%their transit time, since the epoch of occultation was not fit as a free
%parameter.'
Finally, \citet{huitson_gemini_2017} acquired optical transmission
spectra with the 8.1-m Gemini South telescope over 2011 to 2014, one
transit per season.  The per-point standard deviation of their
lightcurves is a few hundred parts per million, within a factor of 3
from the photon noise limit.  The uncertainties on their reported
transit times~--~on average 5.6~seconds~--~ are small, but possible
given the quality of their data~\citep[cf.][]{carter_analytic_2008}.

% There have been other studies of WASP-4b for which the transit timing
% data are either hetereogenous, not available, or not useable.  We
% omitted the timing measurements from
% \citet{southworth_high-precision_2009}, since there were technical
% problems with the computer clock at the time of
% observation~\citep{nikolov_wasp-4b_2012}.  The timing study
% by~\citet{petrucci_no_2013} ruled out transit timing variations larger
% than 54~seconds over the previous 5 years, but did not report transit
% times.  \citet{baluev_benchmarking_2015} collected transit times for
% many hot Jupiters, including WASP-4b, and included times from the
% Exoplanet Transit Database
% (ETD)\footnote{\url{http://var2.astro.cz/ETD/}}
% \citep{poddany_ETD_2010}.  Since the ETD data come from heterogeneous
% sources, their timestamps are less clearly documented, and the times
% are thus more prone to systematic errors, we omit them from
% consideration.
% %We consider the effects of including ETD data in
% %Appendix~\ref{sec:verify_archival_times}.
% \citet{may_mopss_2018} obtained optical spectra using IMCAS on
% Magellan, and found a flat transmission spectrum, in agreement with
% the results reported by \citet{huitson_gemini_2017}.  They did not
% report transit times.

There are fewer available occultation times.
\citet{beerer_secondary_2011} observed two occultations of WASP-4b using
warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
\citet{caceres_ground-based_2011} detected an occultation from the
ground in the ${\rm K}_{\rm s}$ band, and gave a time in HJD, without
specifying the time standard.  We correspondingly added $69.184$ seconds
of uncertainty to their reported errors, in quadrature.
%Finally, as
%previously mentioned, \citet{ranjan_atmospheric_2014} acquired
%occultation data with HST but kept the epoch as a fixed parameter.
\citet{zhou_secondary_2015} acquired
occultation data from the Anglo-Australian Telescope, focusing on the
eclipse depth, but also fitting for $e\cos\omega$.  We converted their
$e\cos\omega$ result into a midtime, using their ephemeris and the
standard formula \citep[{\it
e.g.},][]{charbonneau_detection_2005,winn_exoplanet_2010}
\begin{equation}
  t_{\rm occ}(E) =
  t_0 +  P E  +
  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
  \label{eq:occultation_time}
\end{equation}
This gives us a total of 4 occultations.  Although the occultation
times are uncertain compared to the transits, we include them for
completeness, and because they help in modeling the timing
variations.

@@

After collecting the literature times and measuring the TESS times, we
compared the observed time to the prediction based solely on the
literature times. The result is shown in
Figure~\ref{fig:arrived_early}. The TESS observations seem to have
arrived early.  One immediate concern was systematic errors in either
the archival or the TESS timestamps. The tests we used to mitigate the
latter possibility are described in Appendix~\ref{sec:verify_tess}.

@@

% Assuming the error bars on the transit midtimes are gaussian and
% independent, we then proceed by fitting each model to the timing data.
% For the linear model, we took uniform priors over small windows around
% the expected parameters.  For the quadratic model we did the same, and
% took the quadratic term to have a uniform prior corresponding to
% $
%   Q_\star' \sim \mathcal{U}[10^3, 10^9],
% $
% for $Q_\star'$ the modified quality factor defined in
% \S~\ref{sec:implications}.  For the precession model, we took the same
% priors in epoch and sidereal period, and drew the remaining parameters
% from wide uniform distributions.

%TODO: final numbers
Figure~\ref{fig:times} shows the residuals with respect to the linear
model.  The best linear fit has $\chi^2 = 149$ and 58 degrees of
freedom.  The best quadratic fit has $\chi^2 = 48.0$ and 57
degrees of freedom, compared to $\chi^2 = 51.0$
and 55 degrees of freedom for the best-fit precession model.

%TODO: final numbers
The difference in $\chi^2$ between the linear and quadratic fit
corresponds to $p \approx 10^{-23}$. %In other words, assuming the linear
%model is true, there is a $10^{-23}$ probability that simply by chance
%we would see data that favors the quadratic model more strongly than the
%observations.
Clearly, it is more likely that some systematic error could be
affecting one of the most important sets of times: either the earliest
times, the Huitson et al.\ Gemini times (epoch numbers 44, 322, 591,
and 851), or the TESS times.  We have taken every precaution against
systematic offsets in the TESS dataset
(Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by their
reported times (Huitson, priv.  comm.).  Since the linear model
provides a poor fit to the data, we discard it from further
consideration.
